Papers_search_solutions

### 2019-02-06 Papers_search.Rmd
typos fixed

In this exercise, we will use a few basic tools to illustrate how to perform the first few steps of a meta-analysis: refining the question, searching the literature, and screening the found studies (these are the same steps as in Systematic Reviews, so the tools are the same and can be usually described as systematic literature/evidence reviewing tools).  
Note that some R packages (and many other online/software tools) are available (and more are bieng developed) that can help with some of the tasks during this exercise, but there is no single "perfect" tool for the whole process usable for all disciplines and types of questions (although some attempts have been made). We simply do not have enough time to introduce more tools and the main purpose of this exercise is to familiarize you with the basic principles/issues that apply when you use any tools. You can always try alternative tools in your own time.

@@

**QUESTION 0:**  
What is your initial feeling about the meta-analytical question that can be derived from the text fragment cited above?

@@

**QUESTION 11:**  
Data type: should we accept any type of data for meta-analysis? (e.g. descriptive statistics, and which ones, or also inferential model statsistics, and which ones? This will be linked to the effect size you are planning to use!)

@@

Full-text screening is more strict and we want to have answers to all our inclusion criteria. In principle we only have "yes" and "no" answers to the screening questions. However, in practice, if a study looks very relevant but some information is ambiguous or missing, you may decide to contact the authors to get some extra information or data. in such case, you can provisionary include such papers and make final decision once you get, or not, the replies.

@@

Normally, we attempt to do a fairly comprehensive literature search (a full systematic review). The reason is that we want to have as many relevant data points as possible to increase the power of statistical analyses. The other reason is that we make effort to capture obscure studies that potentially are more likely to report non-significant results, and ttherefore help reduce publication bias in our data set. Search strategy should include using at least two wide-coverage/interdisciplinary databases (we recommend Scopus and Web of Science), and a few additional sources of literature such as forward and backward reference searching (looking at cited and citing papers), publication lists of the key authors or labs, searching for unpublished studies and data sets (we will not cover these here).

@@

 1. Open Rayyan (we assume you already created a free account) and log in.    
 2. My Reviews -> New REVIEW...   
 3. Enter project title, e.g. "Terminal Investment"" and description, e.g. "test"."   
 4. Select files -> upload *Scopus10.ris* and press Continue.     
It may take several minutes for larger files and slower connections, but eventually the uploaded references should appear in the left panel under "Search methods" (you can upload more files with references via "Add new" link). Click on the "Uploaded References [scopus10.ris]" link in this panel to see the references in the right panel. Generally clicking on the links within panels on the left will filter the records in the right panels to include only highlighted sets (to get rid of the filter click on it again to remove highlighting).  
In the left panels there are several sub-panels.  
 5. In the "Keywords for include"" you can add (via "Add new" link) following terms (or at least a few of these): challenge, challenged, injected, injection, infected, infection, PHA, SRBC, vaccine (these terms will be highlighted green in the titles and abstracts to help them find more easily; Rayyan tries to populate this list by itself - but its algorithm is medical-centered so you should adjust this list for your projects).   
 6. "Keywords for exclude" - " set to: review, reviewed, model, human, people, women, men (these terms will be highlighted red in the titles and abstracts to help them find more easily; again, Rayyan tries to populate this list by itself).    
 7. Click on "All reviews" button in the top right corner to go to the project lists, then open your project again by choosing it from the list. A view with a pie chart (your screening progress) and a few buttons will appear. If you need to invite other people to do screening in parallel with you, do that by clicking on the "Invite" button. Stick with "Collaborator" selection and enter the email and message (e.g. "help me!"). Open your project again - you should see your collaborators added with some info on their work so far. Press on "Show" button to see the references view.  
 8. Click on the first review from the top - an abstract and other record info will appear in the bottom panel.  
 9. Using our abstract screening decision tree decide whether include this paper for the next stage of screening and press the "Include" or "Exclude" button accordingly. And do the assessment for the remaining records. As you go you can create and assign new labels to individual records, e.g.: "unsure/check". You could also enter reasons for exclusion, e.g. "human" (we normally don't do this as there are often hundreds or thousands of excluded records per project at this stage).  
10. Go to "All reviews" and check if other people finished their screening, if applicable. After that you can lift the "blind" mode (when colleborators cannot see each others decisions). For collaborative screens Rayyan will now provide filters to display all consensus included, consensus excluded and records with conflicting decisions. If you have any "conflicts" have a look again and discuss why you couldn't agree on a given record , come to agreement and make final decision. Export the records with final consensus decisions, for documentation, into a .csv file. You can filter the records into “Included” and just export those, also you can export them as a .bib or .ris file into a reference manager where you will collect full texts for the enxt stage of the screening.

>>>>

In this exercise, we will use a few basic tools to illustrate how to perform the first few steps of a meta-analysis: refining the question, searching the literature, and screening the found studies (these are the same steps as in Systematic Reviews, so the tools are the same and can be usually described as systematic literature/evidence reviewing tools).  
Note that some R packages (and many other online/software tools) are available (and more are being developed) that can help with some of the tasks during this exercise, but there is no single "perfect" tool for the whole process usable for all disciplines and types of questions (although some attempts have been made). We simply do not have enough time to introduce more tools and the main purpose of this exercise is to familiarize you with the basic principles/issues that apply when you use any tools. You can always try alternative tools in your own time.

@@

**QUESTION 0:**  
What is your initial feeling about the meta-analytic question that can be derived from the text fragment cited above?

@@

**QUESTION 11:**  
Data type: should we accept any type of data for meta-analysis? (e.g. descriptive statistics, and which ones, or also inferential statistics, and which ones? This will be linked to the effect size you are planning to use!)

@@

Full-text screening is more strict and we want to have answers to all our inclusion criteria. In principle we only have "yes" and "no" answers to the screening questions. However, in practice, if a study looks very relevant but some information is ambiguous or missing, you may decide to contact the authors to get some extra information or data. in such case, you can provisionally include such papers and make final decision once you get, or not, the replies.

@@

Normally, we attempt to do a fairly comprehensive literature search (a full systematic review). The reason is that we want to have as many relevant data points as possible to increase the power of statistical analyses. The other reason is that we make effort to capture obscure studies that potentially are more likely to report non-significant results, and therefore help reduce publication bias in our data set. Search strategy should include using at least two wide-coverage/interdisciplinary databases (we recommend Scopus and Web of Science), and a few additional sources of literature such as forward and backward reference searching (looking at cited and citing papers), publication lists of the key authors or labs, searching for unpublished studies and data sets (we will not cover these here).

@@

 1. Open Rayyan (we assume you already created a free account) and log in.    
 2. My Reviews -> New REVIEW...   
 3. Enter project title, e.g. "Terminal Investment"" and description, e.g. "test"."   
 4. Select files -> upload *Scopus10.ris* and press Continue.     
It may take several minutes for larger files and slower connections, but eventually the uploaded references should appear in the left panel under "Search methods" (you can upload more files with references via "Add new" link). Click on the "Uploaded References [scopus10.ris]" link in this panel to see the references in the right panel. Generally clicking on the links within panels on the left will filter the records in the right panels to include only highlighted sets (to get rid of the filter click on it again to remove highlighting).  
In the left panels there are several sub-panels.  
 5. In the "Keywords for include"" you can add (via "Add new" link) following terms (or at least a few of these): challenge, challenged, injected, injection, infected, infection, PHA, SRBC, vaccine (these terms will be highlighted green in the titles and abstracts to help them find more easily; Rayyan tries to populate this list by itself - but its algorithm is medical-centered so you should adjust this list for your projects).   
 6. "Keywords for exclude" - " set to: review, reviewed, model, human, people, women, men (these terms will be highlighted red in the titles and abstracts to help them find more easily; again, Rayyan tries to populate this list by itself).    
 7. Click on "All reviews" button in the top right corner to go to the project lists, then open your project again by choosing it from the list. A view with a pie chart (your screening progress) and a few buttons will appear. If you need to invite other people to do screening in parallel with you, do that by clicking on the "Invite" button. Stick with "Collaborator" selection and enter the email and message (e.g. "help me!"). Open your project again - you should see your collaborators added with some info on their work so far. Press on "Show" button to see the references view.  
 8. Click on the first review from the top - an abstract and other record info will appear in the bottom panel.  
 9. Using our abstract screening decision tree decide whether include this paper for the next stage of screening and press the "Include" or "Exclude" button accordingly. And do the assessment for the remaining records. As you go you can create and assign new labels to individual records, e.g.: "unsure/check". You could also enter reasons for exclusion, e.g. "human" (we normally don't do this as there are often hundreds or thousands of excluded records per project at this stage).  
10. Go to "All reviews" and check if other people finished their screening, if applicable. After that you can lift the "blind" mode (when collaborators cannot see each others decisions). For collaborative screens Rayyan will now provide filters to display all consensus included, consensus excluded and records with conflicting decisions. If you have any "conflicts" have a look again and discuss why you couldn't agree on a given record , come to agreement and make final decision. Export the records with final consensus decisions, for documentation, into a .csv file. You can filter the records into “Included” and just export those, also you can export them as a .bib or .ris file into a reference manager where you will collect full texts for the nest stage of the screening.

### 2019-02-06 Papers_search_solutions.Rmd
typos fixed

**ANSWER 11:**   
It depends on the effect size we are planning to use. In our project it can be Cohen's d (Hedges g) or lnRR, which are typically calculated from mean, SD or SE and N (sample sizes). However, they can be also inferred from some of the test statistics, or even proportions, if means and other descriptive statistics are not available. Other considerations include what to do if some data is missing or unclear.

@@

**ANSWER 15:**    
The question is "Is the full text of this study available for our assessment?"   
It is often not explicitly stated, but without the full text it is impossible to extract all the necessary information about the study and the data. 
It is a good practice to keep this in mind and use as one of your potential inclusion/exclusion critera.

**ANSWER 16:**   
We think 4 should have been excluded: Simmons2009 and Clancy 2013 as human studies, Gershman2010 and Gonzales-Tokman 2013 as not likely to have any reproduction-related measurements.

>>>>

**ANSWER 11:**   
It depends on the effect size we are planning to use. In our project it can be Cohen's d (Hedges g) or lnRR, which are typically calculated from mean, SD or SE and N (sample sizes). However, Cohen's d (Hedges g) can be also inferred from some of the test statistics, or even proportions, if means and other descriptive statistics are not available. But one cannot get lnRR from inhferential statistics. Other considerations include what to do if some data is missing or unclear.

@@

**ANSWER 15:**    
The question is "Is the full text of this study available for our assessment?"   
It is often not explicitly stated, but without the full text it is impossible to extract all the necessary information about the study and the data. 
It is a good practice to keep this in mind and use as one of your potential inclusion/exclusion criteria.

**ANSWER 16:**   
We think 4 should have been excluded: Simmons2009 and Clancy2013 as human studies, Gershman2010 and Gonzales-Tokman2013 as not likely to have any reproduction-related measurements.
